### 2013-02-04 paper.tex
Paper tex added

%% LyX 2.0.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage{amsmath}
\usepackage{amssymb}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{babel}
\usepackage{fancyhdr}
\usepackage{lastpage}

\makeatother

\usepackage{babel}
\begin{document}
\pagestyle{fancy} \lhead{Team \# 23813} \rhead{Page \thepage\ of
\pageref{LastPage}}

\makeatletter \let\@oddfoot\@empty \let\@evenfoot\@empty \makeatother

\title{title}

\author{Control Number: 23813}

\date{\today}
\maketitle
\begin{abstract}
\thispagestyle{empty}

In response to the potential of water shortages in the United States
by 2025, we present a series of recommended infrastructure developments
intended to most efficiently meet future demands. We first create
a projection of future water supply and demand at the county level
based on extrapolating past data collected by the US Geological Survey,
providing a visualization of water needs across the country. Based
on these projections, and geographical data concerning the location
of each county, we determine that a water management strategy primarily
based on the construction of a pipeline network is ideal. We model
counties as vertices and pipelines as edges in order to employ a minimum
spanning tree algorithm to determine the optimal placement of new
pipelines in order to minimize costs. Anticipating the possibility
of funding being less than that required to construct all recommended
infrastructure, we create a priority ordering of the pipelines, using
a utility function based on county population data. We also consider
the water supply of the country as a whole, providing recommendations
as to the number of de-salinization plants required to meet the change
in demand.

\newpage{}

\thispagestyle{empty}

To the United States Government:

$\:$

In light of the 2012-2013 drought season, and the possibility of potentially
more severe droughts in the future, our department has conducted an
analysis of how best to satisfy the water needs of the United States.

For the full details of our findings, please refer to our full paper,
attached below.

\newpage{}

\tableofcontents{}

\newpage{} 
\end{abstract}

\section{Introduction}

The availability of freshwater is growing increasingly severe as a
limiting factor in world development. In particular, issues of pollution,
climate change, and the possibility of increased water demand in the
future could potentially result in more frequent water shortages in
some parts of the United States.

It is projected that by 2025 the world's total demand for water will
exceed available supplies \cite{key-1}. In response to growing concerns
regarding water management, we construct a water management plan for
the United States in 2013, with the goal of developing infrastructure
to meet projected needs by 2025. We will provide recommendations as
to where new infrastructure ought to be placed in order to most efficiently
alleviate water shortages, based on a series of projections and a
graph-based model focused on constructing a pipeline network.

\subsection{Overview of Our Approach}

We begin our paper with a projection of the likely levels of water
supply (given no additional interference) and demand in the United
States by 2025 on a county-by-county basis. Our primary source of
data is the County-level water use data made publically available
by the US Geological Survey (USGS) in years 1985, 1990, 1995, 2000,
and 2005 \cite{key-2,key-3,key-4,key-5,key-6}. Using our projected
figures, we describe a means of minimizing the expected cost of constructing
water storage and water purification systems using a network flow
algorithm.

First, we project the County-level water needs in 2025 using water
use data from the years listed, making the assumption that water use
equals the current water demands.

This paper attempts to address key aspects of a feasible water management
strategy, based on the available data. With our projection of expected
water needs in mind, we aim to meet demands by investing in the following
infrastructure: 
\begin{itemize}
\item De-salinization plants 
\item Water purification facilities 
\item Reservoirs and dams
\end{itemize}
However, it is not sufficient to simply have more water available
in the country's supplies as a whole. It's also critical that each
individual location be able to access that water. For this purpose
we aim to find the optimal locations to place pipelines.

This paper is primarily concerned with the construction of new water
supply infrastructure. Thus, there are some aspects of the water management
process that will not be further considered: 
\begin{itemize}
\item Public education: As with all issues of resource conservation, the
behaviour of the general public has an effect on water supplies. This
will not be considered here, for several reasons. The key issue is
that there is a lack of empirical evidence suggesting that currently
known public education strategies will work to minimise water usage
in the United States (developing such strategies further is beyond
the scope of this paper). The United States in particular has been
particularly resistant to attempts to reduce per capita water consumption\cite{key-7}.
More generally, even water education programmes considered particularly
successful, such as that of Singapore, have not managed to make a
huge impact on consumption, with an approximate 5 per cent reduction
in household water use over 6 years\cite{key-8}. Finally, the effect
of water education programmes is likely to have a similar effect in
each location. This means that areas that would be most in need of
additional infrastructure taking into account such programmes would
be essentially the same as those areas most in need while not taking
into account such programmes. 
\item Maintenance costs of current infrastructure: minimizing maintenance
costs depends on technological advancement and management, but is
essentially a fixed cost irrespective of the new infrastructure we
decide to implement.
\item Cost of obtaining data: Much of our analysis depends on USGS data.
We note that the cost of obtaining this survey data is negligible
(the WaterSMART initiative, which collects much of the data, cost
\$21 million this year\cite{key-9}. In comparison, de-salinization
facilities can easily cost billions to construct\cite{key-10}) compared
to the utility of superior decision making in constructing water-related
infrastructure (among various other things the data is useful for).
\end{itemize}

\section{Modeling Infrastructure}

\subsection{Assumptions}

Due to the scale and complexity of the task of water management in
the United States, we make several simplifying assumptions in order
to obtain a workable model.
\begin{itemize}
\item Counties are thought of as discrete entities, occupying points in
space located at their centroid. This assumption enables us to think
of counties as vertices (as in a graph). This assumption can be made
without losing too much accuracy for several reasons: The United States
is split into more than 3,000 counties, with an average area of less
than 3,300 square kilometers, and average length across of less than
60 kilometers. The level of inaccuracy is sufficiently low that it
is workable (in contrast, states for instance are far too large to
be accurately modeled as points).
\item When considering the costs of pipelines, only the distance of pipeline
is considered. The landscape of a region may in practice affect the
length of pipeline required, or may result in additional cost per
length of pipeline, but we shall not model this, as we could not find
data on the difficulty of pipeline construction across different types
of terrain.
\item Pipelines are thought of as bi-directional. This assumption essentially
reflects the reality of pipelines. There are a few exceptional cases
where it is significantly more difficult to send water in one direction,
but these are rare enough that they can be disregarded.
\item Pipelines have variable rates of water transmission, depending on
factors such as size. However, pipelines typically 
\item Some pipelines exist
\end{itemize}

\subsection{Costs\label{sub:Costs}}

The prices per unit for water-related infrastructure are important
constants within our models.

We assume a constant cost per length of pipeline. Estimates for the
cost of pipeline range from approximately \$1.5 million per mile to
\$5 million per mile\cite{key-11}, with the higher figures generally
corresponding to older pipelines constructed less efficiently than
is possible today. We assume a figure of \$2 million per mile for
the purporses of our pricing estimates.

The cost of a de-salinization facility in 2013 is approximately \$14
per gallon per day\cite{key-12}.

\section{Projecting Water Demand}

To inform our water management decisions we extrapolated from the
available data from the years 1985-2005 to project the total water
usage at the county level in 2025. The available data comes in a matrix
$WU[i,j]$ with $i\in\{1985,1990,1995,2000,2005\}$ and $j$ ranging
through the US counties (indexed by FIPS codes), of total water used
per county. In the USGS data \cite{key-2,key-3,key-4,key-5,key-6}
certain peculiarities and missing data arises, for instance:
\begin{itemize}
\item Eight full states are missing water usage data in 2000 altogether.
\item Anomalous data an order of magnitude off arises commonly in the data
in year 1985.
\item Other discrepancies appear in certain counties - for example, a handful
of county FIPS codes go out of existence between 1990 and 1995, and
two new counties are introduced in between 2000 and 2005.
\end{itemize}
A second difficulty with projection is the sheer temporal distance
between the data we have, between 1985 and 2005, and the date at which
we need to model, 2025. This massive extrapolation distance means
that our projections will be highly sensitive to small changes in
model, which requires that we use sensitive outlier analysis before
applying regression analysis on the water usage data.

We apply the following approach to remove outliers from $WU[i,j]$
and project the usage, which we assume to be equal to the total demand,
on the county level in 2025.

\subsection{Total Supply}

We model the overall water requirements of the United States by applying
a linear regression to the total amount of water used across the country
as a whole.

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Year & 1985 & 1990 & 1995 & 2005 & 2013 (proj.) & 2025 (proj.)\tabularnewline
\hline 
Total Water Use (Bgal/day) & 398465 & 407980 & 401554 & 410465 & 413695 & 419355\tabularnewline
\hline 
\end{tabular}

This model

\section{Optimising Placement of Pipelines}

Our modelling of future water needs is consistent with external research
on drought locations\cite{key-13}, in that it suggests many of the
areas at risk of drought do not have the proximity to water sources
required to make the construction of de-salinization plants feasible.

The droughts in central America tend to occur in periods of minimal
rainfall, and as the surrounding rivers and groundwater aquifers also
dry up.

This suggests that there is no currently known way to locally construct
infrastructure to alleviate the water shortages in many of the areas
in which water is most needed.

With this fact in mind, along with the knowledge that some areas in
the United States have a surplus of water, we consider a plan implemented
elsewhere in the world in a similar situation: a grid of pipelines
(aqueducts), as implemented in Queensland, Australia\cite{key-14},
in order to move water from counties with surplus water to those suffering
drought conditions.

\subsection{Modeling a Grid of Pipelines}

To optimise the placement of pipelines, we parse geographical data
for each of the counties, and produce a graph in which counties are
represented as vertices, and undirected edges are placed between adjacent
counties, with weights corresponding to the distance between the counties
being connected.

We aim to connect all of the mainland counties (excluding Alaska and
Hawaii, which our model predicts to be in little need of additional
water infrastructure) while minimizing the length of pipeline used,
and hence minimizing the cost of the project.

To do this, we implement a variant of a minimum spannning tree algorithm
on the complete graph of counties with edges weighted by distance:

$\:$

(add a flowchart here)

$\:$

The end result is a spanning tree of minimal total length connecting
all counties of the US together. This massive project would resolve
all local fluctuations in water availability, allowing high-volume
water transport between nearby counties and from distance regions
to meet seasonal needs in certain areas.

\subsection{Projected Cost of the Grid}

To estimate the cost of building a national high-volume aqueduct system
from scratch, we estimated in Section \ref{sub:Costs} using publically
available figures that the average cost per mile is approximately
\$2 million per mile as technology becomes cheaper and more efficient.

\section{The Case of Insufficient Funding}

Our model projects levels of supply and demand in 2025, and proceeds
to recommend infrastructure to meet that demand while minimising cost.
This reflects the main precedent in major development of water supply
infrastructure in the United States, namely the \$24.6 billion Clean
Water Act \cite{key-15}, established with the aim of eliminating
all pollution of water supplies.

However, it is possible there will be insufficient funding for this
project. In this case, it will not be possible to construct all of
the recommended infrastructure. We approach this problem by creating
a priority listing, ordering the suggested developments in terms of
utility per dollar. To this end, we employ the following formula:

\section{Sensitivity and Variance}

There are two anticipated causes of significant variance in water
levels:
\begin{itemize}
\item Pathological rainfall levels (ie. extended periods of drought) 
\item Significant unexpected change in population or behaviour 
\end{itemize}
Water systems in general are resistant to small variance in water
supply or demand, and our system in particular fares well in this
regard. Many of our suggested developments are particularly well equipped
to handle over-supply: our model recommends the construction of additional
pipelines and de-salinization plants — pipeline flow can be easily
reduced, while de-salinization plants can be set to operate at partial
capacity, saving electricity and avoiding over-supply.

Slight shortages of water in the United States have consistently been
dealt with successfully by implementing targeted water restrictions
\cite{key-16}. There is little argument that the limitations, largely
on gardening water, do not have a huge impact on people.

\section{Other Impacts}

The core of our water management strategy is a large network of pipelines
across the country in order to shift water to locations where it is
most needed. The main reason for this is that similar approaches have
proven effective when dealing with issues of drought in inland regions
where groundwater and surface water supplies approach depletion. However,
our plan of relying mostly on pipelines also has some other benefits
over relying on other kinds of infrastructure.

While de-salinization is simply not an option at all for much of central
America, one idea we considered was the construction of additional
dams and reservoirs. One considerable environmental issue associated
with large dams is that they can cause unanticipated changes in overall
river flow\cite{key-17}, totalling potentially huge amounts in damages.

A secondary effect of high national connectivity is decreased reliance
on groundwater.

\section{Conclusions}

Our analyses allow us to make several important conclusions and recommendations
regarding the future of water in the United States. By projecting
future increases in water, we observe that while some parts of the
United States (particularly on the coasts) are not at significant
risk of water shortage by 2025, other parts of the country, particularly
Central America, which already experienced drought in 2012-2013, are
predicted to increase their demands for water in the upcoming years,
and hence will be at risk of \textbf{severe drought}. Moreover, the
area of the country potentially at risk is large, and a significant
portion of the population could be affected. We conclude that there
is significant utility in investing further in water infrastructure
in the United States.

To an extent

The cost of the proposed infrastructure developments are significant,
but must be placed in context. The cost of the 2012-2013 drought season
on the United States economy is estimated at between \$75 billion
and \$150 billion, and the potential total cost to the economy of
drought-related occurrences between 2013 and 2025 is much greater
even still.

As a result, we firmly recommend that the entirety of our proprosed
infrastructure developments, particularly the national pipeline network,
be implemented over the upcoming years.
\begin{thebibliography}{10}
\bibitem{key-1}Cai, X., Cline, S., Rosegrant, M. (2002). Global Water
Outlook to 2025. \emph{International Food Policy Research Institute.}

\bibitem[2]{key-2}Estimated Use of Water in the United States County-Level
Data for 1985 (1955). \emph{US Geological Survey. }

\bibitem[3]{key-3}Estimated Use of Water in the United States County-Level
Data for 1990 (1990). \emph{US Geological Survey. }

\bibitem[4]{key-4}Estimated Use of Water in the United States County-Level
Data for 1995 (1995). \emph{US Geological Survey. }

\bibitem[5]{key-5}Estimated Use of Water in the United States County-Level
Data for 2000 (2000). \emph{US Geological Survey.}

\bibitem[6]{key-6}Estimated Use of Water in the United States County-Level
Data for 2005 (2005). \emph{US Geological Survey.}

\bibitem[7]{key-7}Per Capita Water Use. Retrieved February 1, 2013,
from http://www.pepps.fsu.edu/safe/pdf/sc1.pdf

\bibitem[8]{key-8}Loh, J. (n.d.). Water management: Learning from
Singapore's water succcess. Retrieved February 1, 2013, from http://workingwithwater.filtsep.com/view/934/water-management-learning-from-singapore-s-water-success/

\bibitem[9]{key-9}2013 USGS Budget Proposal. Retrieved February 1,
2013, from http://www.usgs.gov/newsroom/article.asp?ID=3103

\bibitem[10]{key-10}Moran, T. (2007). Water Wars: Quenching Las Vegas'
Thirst. Retrieved February 1, 2013, from http://abcnews.go.com/Nightline/story?id=3012250

\bibitem[11]{key-11}Mission 2012: Clean Water. Retrieved February
4, 2013, from http://web.mit.edu/12.000/www/m2012/finalwebsite/solution/glaciers.shtml

\bibitem[12]{key-12}City of Carlsbad Seawater Desalination, Retrieved
February 4, 2013, from http://www.carlsbadca.gov/services/departments/water/pages/seawaterdist.aspx

\bibitem[13]{key-13}US Drought Monitor. Retrieved February 3, 2013,
from http://droughtmonitor.unl.edu/

\bibitem[14]{key-14}South East Queensland water grid. Retrieved Februrary
3, 2013, from http://www.dsdip.qld.gov.au/infrastructure-delivery/south-east-queensland-water-grid.html

\bibitem[15]{key-15}Burian, S., Nix, S., Pitt, R., Durrans, S. (2000).
Urban Wastewater Management in the United States: Past, Present and
Future. Retrieved February 1, 2013, from http://www.sewerhistory.org/articles/whregion/urban\_wwm\_mgmt/urban\_wwm\_mgmt.pdf

\bibitem[16]{key-16}Kennedy, D., Klein, R., Clark, M. (2004) Use
and Effectiveness of Municipal Water Restrictions During Drought in
Colorado. Retrieved February 1, 2013, from http://sciencepolicy.colorado.edu/admin/publication\_files/resource-296-water\_restrictions\_jawra.pdf

\bibitem[17]{key-17}China's Three Gorges Dam. Retrieved February
3, 2013, from https://www.mtholyoke.edu/\textasciitilde{}vanti20m/classweb/website/environmentalimpact.html

\bibitem[18]{key-18}http://www.wunderground.com/blog/JeffMasters/article.html?entrynum=2326\end{thebibliography}

\end{document}